---
title: "Communication"
---

## Notes

labeling plots text

```{r}
library(tidyverse)
label_info <- mpg |>
  group_by(drv) |>
  arrange(desc(displ)) |>
  slice_head(n = 1) |>
  mutate(
    drive_type = case_when(
      drv == "f" ~ "front-wheel drive",
      drv == "r" ~ "rear-wheel drive",
      drv == "4" ~ "4-wheel drive"
    )
  ) |>
  select(displ, hwy, drv, drive_type)

label_info
#> # A tibble: 3 × 4
#> # Groups:   drv [3]
#>   displ   hwy drv   drive_type       
#>   <dbl> <int> <chr> <chr>            
#> 1   6.5    17 4     4-wheel drive    
#> 2   5.3    25 f     front-wheel drive
#> 3   7      24 r     rear-wheel drive
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  geom_text(
    data = label_info, 
    aes(x = displ, y = hwy, label = drive_type),
    fontface = "bold", size = 5, hjust = "right", vjust = "bottom"
  ) +
  theme(legend.position = "none")
#> `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
```

nudge and rectangle around label

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  geom_label(
    data = label_info, 
    aes(x = displ, y = hwy, label = drive_type),
    fontface = "bold", size = 5, hjust = "right", alpha = 0.5, nudge_y = 2,
  ) +
  theme(legend.position = "none")
#> `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
```

geom_label_repel ggrepel package adjust labels for you

```{r}
library(ggrepel)
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  geom_label_repel(
    data = label_info, 
    aes(x = displ, y = hwy, label = drive_type),
    fontface = "bold", size = 5, nudge_y = 2,
  ) +
  theme(legend.position = "none")
#> `geom_smooth()` using method = 'loess' and formula = 'y ~ x'
```

great use of information

```{r}
potential_outliers <- mpg |>
  filter(hwy > 40 | (hwy > 20 & displ > 5))
  
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_text_repel(data = potential_outliers, aes(label = model)) +
  geom_point(data = potential_outliers, color = "red") +
  geom_point(data = potential_outliers, color = "red", size = 3, shape = "circle open")
```

i think you have to create a data frame everytime

```{r}
label_info <- mpg |>
  summarize(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \nrelated to decreasing fuel economy."
  )

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_text(
    data = label_info, aes(label = label), 
    vjust = "top", hjust = "right"
  )
```

## Questions

## Solutions

prerequisites

```{r}
library(tidyverse)
library(ggrepel)
library(patchwork)

```

## Exercises 12.2.1

1.  Create one plot on the fuel economy data with customized `title`, `subtitle`, `caption`, `x`, `y`, and `color` labels.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs( title = "insert title here",
        subtitle = "insert subtitle here",
        caption = "insert source here",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    color = "Car type"
  )
```

2.Recreate the following plot using the fuel economy data. Note that both the colors and shapes of points vary by type of drive train

```{r}
glimpse(mpg)

mpg %>% ggplot(aes(x = cty,y = hwy,color = drv,shape = drv)) +
  geom_point() +
  labs(x = "City MPG",
        y = "Highway MPG",
        shape = "Type of drive train",
       color = "Type of drive train"
       )
```
