---
title: "Chapter 12 Exploratory data analysis"
format: html
editor: visual
---

# Notes

coord_cartesian() zoom into y

```{r}
ggplot(diamonds, aes(x = y)) + 
  geom_histogram(binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))
```

# Questions

# Solutions

## Exercise 12.3.3

1.Explore the distribution of each of the `x`, `y`, and `z` variables in `diamonds`. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
?diamonds

ggplot(diamonds, aes(x = x , y = y)) + geom_point()
```

x is the length, y is the width, z is the depth

2.Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the `binwidth` and make sure you try a wide range of values.)

```{r}
ggplot(diamonds, aes(x = price)) + geom_histogram(binwidth = 2000)
```

3.How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = .1)
```
